### 2020-12-17 paper/sex_selection_spacing-ver5.tex
Edit Abstract

\noindent India has experienced an increasingly male-biased sex ratio at birth following
the introduction of prenatal sex determination techniques. 
I examine three questions.
First, how did birth spacing respond to this and other secular developments? 
Second, did changes in birth spacing make us overestimate the decline in cohort fertility? 
Finally, did the advent of sex-selective abortions change the relationship between infant mortality and
birth spacing? 
Well-educated Hindu women with no sons had the most substantial lengthening of birth 
intervals and the most male-biased sex ratios, likely from sex-selective abortions. 
Hence, some women with no sons now have longer birth intervals than those with
sons, reversing the traditional spacing pattern. 
Less-educated women continued short spacing when they have girls, with only limited 
evidence of male-biased sex ratios. 
Because of the rapid lengthening in spacing, period fertility rates substantially 
overestimated how fast cohort fertility fell. 
Furthermore, predicted cohort fertility remains above replacement for all but the 
best-educated urban women. 
If lengthening of birth intervals occurs from repeated abortions, the associated short 
pregnancy spacing may counteract any positive effects of longer birth spacing. 
There is no evidence of this effect on infant mortality. 
Finally, judging from sex ratios, sex-selective abortion use is not declining.

>>>>

\noindent India has experienced an increasingly male-biased sex ratio at birth following
the introduction of prenatal sex determination techniques. 
I examine three questions.
First, how did birth spacing respond to this and other secular developments? 
Second, did changes in birth spacing make us overestimate the decline in cohort fertility? 
Finally, did the advent of sex-selective abortions change the relationship between birth 
spacing and infant mortality? 
Well-educated Hindu women with no sons had the most substantial lengthening of birth 
intervals and the most male-biased sex ratios, likely from sex-selective abortions. 
Consequently, some women with no sons now have longer birth intervals than those with sons, 
reversing India's traditional spacing pattern.
Less-educated women continued short spacing when they have girls, with only limited
evidence of male-biased sex ratios. 
Because of the overall lengthening in spacing, period fertility rates substantially 
overestimated how fast cohort fertility fell. 
Furthermore, predicted cohort fertility remains above replacement for all but the 
best-educated urban women. 
If lengthening of birth intervals occurs from repeated abortions, the associated
short pregnancy spacing may counteract any positive effects of longer birth spacing. 
There is no evidence of this effect on infant mortality. 
Finally, judging from sex ratios, sex-selective abortion use is not declining.
